Ground state energy of quantum dots using the coupled cluster method

### 2018-06-13 project_2/article/article.tex
Elaborate on the HF-method
Link to example script on Github

            The neatest, yet arguably the most abstract, way to write the
            Hartree-Fock equations for electron $i$ is
            \begin{equation}
                f \varphi_p = \varepsilon_p \varphi_p,
            \end{equation}
            where $f$ is the Fock operator, $\varphi_p$ are eigenstates of the
            Fock operator consisting of a set of one-electron wave functions
            called the Hartree-Fock molecular orbitals, and $\varepsilon_p$ are
            the eigenenergies of the Fock operator. The Fock operator, in matrix
            notation, is given by
            \begin{equation}
                f^p_q = h^p_q + J(D)^p_q - \half K(D)^p_q,
                %F_{pq} = H_{pq} +  J(D)_{pq} - \frac{1}{2}K(D)_{pq},
            \end{equation}
            where $h^p_q$ is the one-body operator, while the two-body operator
            is divided into what we call a direct part,
            \begin{equation}
                J(D)^p_q = \bra{pr}u\ket{qs}D_{sr},
            \end{equation}
            and an exchange part,
            \begin{equation}
                K(D)^p_q = \bra{pr}u\ket{sq}D_{sr}.
            \end{equation}
            Pay special attention to the non-antisymmetric two-body elements in
            the direct and exchange part.  The direct part of the two-body
            operator is comparable to classical Coloumb repulsion, while the
            exchange does not have a classical analog as it arises from the
            antisymmetry requirement of the wavefunction. The matrix $D_{sr}$ is
            the density matrix of the system and will be defined shortly.

            Introducing a basis set transforms the Hartree-Fock equations into
            the Rothaan equations
            \begin{equation}
                FC = SC\varepsilon.
            \end{equation}
            This is a generalised eigenvalue problem where
            \begin{align}
                S^{p}_{q} = \braket{p}{q},
            \end{align}
            i.e., the \emph{overlap matrix}. We now define the density matrix
            $D$ by
            \begin{align}
                D^{p}_{q} \equiv C^p_{\alpha}{C^{\alpha}_q}^{*},
            \end{align}
            Since the Fock matrix $F$ depends on it's own solution through the
            orbitals, the eigenvalue problem must be solved
            iteratively\footnote{This is also the reason why the Roothan-Hall
            equations are often called the self-consistent-field procedure.}

@@

            The transformations necessary to make all these variations will now
            be contained in $C$. So if we start from a harmonic oscillator (HO)
            basis, $\ket{\alpha}$ we can transform into a better basis, called a
            Hartree-Fock (HF) basis, $\ket{p}$.
            \begin{equation}
                \ket{p} = C^p_\alpha\ket{\alpha}.
            \end{equation}
            We will see how the coupled cluster method with double excitations
            works with both the HO basis and the HF basis, but not equally well.

\section{Implementation}
    We have costructed a flexible framework for making performing Hartree-Fock
    self-consistednt field (SCF) and coupled cluster with double excitations (CCD)
    computations. The entire code base consists of package for \lstinline{python}
    that is easy to install globally on any computer.

    Because we have employed a mixture of Python and C++ with Cython interfaces,
    the code sturcture may be difficult to undertand for someone without Cython
    experience, but we have tried to make the actual usage of the software as
    painless as possible.

    Most of the computations that are done in a coupled cluster algorithm
    is linear algebra and we found that NumPy, relying on
    BLAS and LAPACK, is most likely as fast as it gets and, at the very least,
    efficient enough for our purposes. The quicker and easier scripting capabilities
    Python provides is worth what little extra speed-up writing the entire
    code base in C++ could have provided.

@@

        The software is easy to install by first cloning the GitHub reopsitory,
        \begin{bash}
            git clone git@github.com:Schoyen/FYS4411.git
        \end{bash}
        Then all you need to do is change directory to the project code folder
        where a \lstinline{MakeFile} is included so you only need to build and
        install,
        \begin{bash}
            cd FYS4411/project_2/coupled-cluster
            make build
            make install
        \end{bash}
        Now everything should be able to run from anywhere on your computer.
        To ensure that all requirements of the program are satified you can
        install with \mintinline{bash}{make installr} instead.

        We have included a sample python script in \autoref{app:sample_script}
        that uses almost all methods and classes in the package, and also compares
        the time each methods takes to fininsh the computations.

@@

\vfill
\clearpage
\newpage
\begin{widetext}
\section{Example Script}

    \label{app:sample_script}

    \vspace{10pt}

    \inputminted[firstline=0, lastline=113]{python}{../scripts/comparison.py}

\end{widetext}

>>>>

            The neatest, yet arguably the most abstract, way to write the
            Hartree-Fock equations for electron $i$ is
            \begin{equation}
                f \varphi_p = \varepsilon_p \varphi_p,
            \end{equation}
            where $f$ is the Fock operator, $\varphi_p$ are eigenstates of the
            Fock operator consisting of a set of one-electron wave functions
            called the Hartree-Fock molecular orbitals, and $\varepsilon_p$ are
            the eigenenergies of the Fock operator. The Fock operator, in matrix
            notation, is given by
            \begin{equation}
                f^p_q = h^p_q + J(D)^p_q - \half K(D)^p_q,
            \end{equation}
            where $h^p_q$ is the one-body operator, while the two-body operator
            is divided into what we call a direct part,
            \begin{equation}
                J(D)^p_q = \bra{pr}u\ket{qs}D_{sr},
            \end{equation}
            and an exchange part,
            \begin{equation}
                K(D)^p_q = \bra{pr}u\ket{sq}D_{sr}.
            \end{equation}
            Pay special attention to the non-antisymmetric two-body elements in
            the direct and exchange part.  The direct part of the two-body
            operator is comparable to classical Coloumb repulsion, while the
            exchange does not have a classical analog as it arises from the
            antisymmetry requirement of the wavefunction. The matrix $D_{sr}$ is
            the density matrix of the system and will be defined shortly.

            Introducing a basis set transforms the Hartree-Fock equations into
            the Rothaan equations
            \begin{equation}
                FC = SC\varepsilon,
            \end{equation}
            where $F$ is the Fock matrix constructed from $f^p_q$.  This is a
            generalised eigenvalue problem where
            \begin{align}
                S^{p}_{q} = \braket{p}{q},
            \end{align}
            i.e., the \emph{overlap matrix}. We now define the density matrix
            $D$ by
            \begin{align}
                D^{p}_{q} \equiv C^p_{\alpha}{C^{\alpha}_q}^{*},
            \end{align}
            Since the Fock matrix $F$ depends on it's own solution through the
            orbitals, the eigenvalue problem must be solved
            iteratively\footnote{This is also the reason why the Roothan-Hall
            equations are often called the self-consistent-field procedure.}

@@

            The transformations necessary to make all these variations will now
            be contained in $C$. So if we start from a harmonic oscillator (HO)
            basis, $\ket{\alpha}$ we can transform into a better basis, called a
            Hartree-Fock (HF) basis, $\ket{p}$.
            \begin{equation}
                \ket{p}
                = \para{\ket{\alpha}\bra{\alpha}}\ket{p}
                = C^p_\alpha\ket{\alpha}.
            \end{equation}
            We will see how the coupled cluster method with double excitations
            works with both the HO basis and the HF basis, but not equally well.

\section{Implementation}
    We have costructed a flexible framework for performing Hartree-Fock
    self-consistent field iterations and coupled cluster with double excitations
    computations. The entire code base consists of package for
    \lstinline{Python} that is easy to install globally on any computer.

    Because we have employed a mixture of Python and C++ with Cython interfaces,
    the code sturcture may be difficult to understand for someone without Cython
    experience, but we have tried to make the actual usage of the software as
    painless as possible.

    Most of the computations that are done in a coupled cluster algorithm is
    linear algebra and we found that NumPy, relying on BLAS and LAPACK, is quite
    fast and at the very least, efficient enough for our purposes. The quicker
    and easier scripting capabilities Python provides is worth what little extra
    speed-up writing the entire code base in C++ could have provided.

@@

        The software is easy to install by first cloning the GitHub reopsitory,
        \begin{bash}
            git clone git@github.com:Schoyen/FYS4411.git
        \end{bash}
        Then all you need to do is change directory to the project code folder
        where a \lstinline{Makefile} is included so you only need to build and
        install,
        \begin{bash}
            cd FYS4411/project_2/coupled-cluster
            make build
            make install
        \end{bash}
        Now everything should be able to run from anywhere on your computer.  To
        ensure that all requirements of the program are satisfied you can
        install with \mintinline{bash}{make installr} instead. Running
        \mintinline{bash}{py.test} on the command-line in the
        top-directory\footnote{That is, in
        \texttt{FYS4411/project\_2/coupled-cluster}.} will run the test scripts
        located in the \texttt{tests}-directory to make sure that the
        installation passed successfully.

        The script that is used to generate most of the data is located in
        \url{https://github.com/Schoyen/FYS4411/blob/master/project_2/scripts/comparison.py}
        and provides a good example on how to use the code base.

@@

%\vfill
\clearpage
%\newpage
%\begin{widetext}
%\section{Example Script}
%
%    \label{app:sample_script}
%
%    \vspace{10pt}
%
%    \inputminted[firstline=0, lastline=113]{python}{../scripts/comparison.py}
%
%\end{widetext}
